Longitudinal Principal Manifold Estimation

### 2023-08-01 paper/main.tex
update simulation results
add adni hippocampus plots
update for new simulations

The embedding functions used to define each manifold are given in Table \ref{table:simulation_embeddings}. In these functions, the value of $\zeta$ and the functional form of $g(\cdot)$ are varied between trials and represent structural change in the underlying manifold over time. The function $g(\cdot)$ may show change with respect to time that is constant, linear, quadratic, or sinusoidal, with $\zeta$ serving as a scalar multiplier. The values of $\iota$ represent within-image noise in the high-dimensional space, and are randomly sampled from a normal distribution with mean zero and variance that is varied between trials. Meanwhile, $\alpha$ and $\beta$ are vectors of length $d$ that are drawn from a normal distribution with mean equal to 1 and variance that is varied between trials. These values describe random fluctuations in the manifold between time points, reflecting noise introduced by different imaging sessions.

@@

A factorial design is used to run the simulation studies, with the factor levels set as follows:
\begin{itemize}
  \item $\alpha, \beta, \zeta \in \left\{0, 0.1, 0.25, 0.5, 1.0, 2.0\right\}$
  \item Study duration: $\left\{1, 2, 5, 10\right\}$
  \item Interval between images: $\left\{0.1, 0.25, 0.5, 1\right\}$
  \item Sample size: $\left\{1000, 10000\right\}$
  \item Longitudinal change model: Constant, Linear, Quadratic, Sinusoidal
\end{itemize}

Thus, each embedding map is simulated a total of 27,648 times. All code required to reproduce the simulation results and the associated figures is provided at \texttt{https://github.com/rjzielinski/lpme-project}.

@@

\begin{table}
  \centering
  \begin{tabular}{|c c c c c c c c c c|}
    \hline
    Method & Case 1 & Case 2 & Case 3 & Case 4 & Case 5 & Case 6 & Case 7 & Case 8 & Case 9 \\
    \hline
    Data & 0.240 (0.346) & 0.489 (0.399) & N/A & 19.2 (48.7) & 0.906 (0.735) & 1.20 (3.16) & N/A & N/A & N/A \\
    LPME & 0.166 (0.668) & 0.324 (0.327) & N/A & 11.8 (31.1) & 0.633 (0.692) & 1.11 (2.61) & N/A & N/A & N/A \\
    PME & 0.265 (0.430) & 0.615 (0.850) & N/A & 19.2 (48.7) & 0.918 (0.831) & 1.26 (3.17) & N/A & N/A & N/A \\
    PC/PS & 0.209 (0.342) & 0.557 (0.319) & N/A & 19.2 (48.7) & 0.863 (0.670) & 0.965 (1.17) & N/A & N/A & N/A \\
    \hline
  \end{tabular}
  \caption{MSD Comparison to True Values, Mean (SD).}
  \label{table:simulation_results_mean}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c c c c c c c c c c|}
    \hline
    Method & Case 1 & Case 2 & Case 3 & Case 4 & Case 5 & Case 6 & Case 7 & Case 8 & Case 9 \\
    \hline
    Data & 0.134 (0.232) & 0.382 (0.649) & N/A & 3.37 (15.7) & 0.729 (1.29) & 0.328 (0.742) & N/A & N/A & N/A \\
    LPME & 0.0661 (0.119) & 0.236 (0.438) & N/A & 3.10 (9.54) & 0.370 (0.889) & 0.364 (0.671) & N/A & N/A & N/A \\
    PME & 0.117 (0.281) & 0.463 (0.668) & N/A & 3.34 (15.6) & 0.675 (1.31) & 0.369 (0.680) & N/A & N/A & N/A \\
    PC/PS & 0.110 (0.214) & 0.471 (0.467) & N/A & 3.33 (15.7) & 0.691 (1.18) & 0.607 (0.329) & N/A & N/A & N/A \\
    \hline
  \end{tabular}
  \caption{MSD Comparison to True Values, Median (IQR).}
  \label{table:simulation_results_median}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{sim_case5}
  \caption{Simulation Case 4}
  \label{fig:sim_case4}
\end{figure}

@@

>>>>

The embedding functions used to define each manifold are given in Table \ref{table:simulation_embeddings}. In these functions, the value of $\zeta$ and the functional form of $g(\cdot)$ are varied between trials and represent structural changes in the underlying manifold over time. The function $g(\cdot)$ may show change with respect to time that is constant, linear, quadratic, or sinusoidal, with $\zeta$ serving as a scalar multiplier. The values of $\iota$ represent within-image noise in the high-dimensional space, and are randomly sampled from a normal distribution with mean zero and variance that is varied between trials. Meanwhile, $\alpha$ and $\beta$ are vectors of length $d$ that are drawn from a normal distribution with mean equal to 1 and variance that is varied between trials. These values describe random fluctuations in the manifold between time points, reflecting noise introduced by different imaging sessions.

@@

A factorial design is used to run the simulation studies, with the factor levels set as follows:
\begin{itemize}
  \item $\alpha, \beta, \zeta \in \left\{0, 0.05, 0.1, 0.25, 0.5, 1.0\right\}$
  \item Study duration: $\left\{1, 2, 5\right\}$
  \item Interval between images: $\left\{0.1, 0.25, 0.5\right\}$
  \item Longitudinal change model: Constant, Linear, Quadratic, Sinusoidal
\end{itemize}

The sample size for each time point is set to 1,000 observations, so each embedding map is simulated a total of 7,776 times. All code required to reproduce the simulation results and the associated figures is provided at \texttt{https://github.com/rjzielinski/lpme-project}.

@@

\begin{table}
  \centering
  \begin{tabular}{|c c c c c c c c c c|}
    \hline
    Method & Case 1 & Case 2 & Case 3 & Case 4 & Case 5 & Case 6 & Case 7 & Case 8 & Case 9 \\
    \hline
    Data & 0.223 (0.311) & 0.524 (0.462) & N/A & 25.5 (75.3) & 0.984 (0.754) & 2.00 (9.29) & N/A & N/A & N/A \\
    LPME & 0.138 (0.174) & 0.363 (0.342) & N/A & 19.8 (69.1) & 0.751 (0.660) & 1.83 (9.19) & N/A & N/A & N/A \\
    PME & 0.231 (0.412) & 0.871 (1.76) & N/A & 25.5 (75.3) & 1.11 (1.36) & 2.02 (9.29) & N/A & N/A & N/A \\
    PC/PS & 0.189 (0.296) & 0.615 (0.357) & N/A & 25.5 (75.3) & 0.928 (0.702) & 1.25 (3.69) & N/A & N/A & N/A \\
    \hline
  \end{tabular}
  \caption{MSD Comparison to True Values, Mean (SD).}
  \label{table:simulation_results_mean}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c c c c c c c c c c|}
    \hline
    Method & Case 1 & Case 2 & Case 3 & Case 4 & Case 5 & Case 6 & Case 7 & Case 8 & Case 9 \\
    \hline
    Data & 0.124 (0.187) & 0.385 (0.640) & N/A & 4.11 (15.9) & 0.939 (1.26) & 0.272 (0.926) & N/A & N/A & N/A \\
    LPME & 0.0707 (0.138) & 0.251 (0.511) & N/A & 3.23 (7.67) & 0.611 (1.05) & 0.290 (0.860) & N/A & N/A & N/A \\
    PME & 0.105 (0.202) & 0.379 (0.708) & N/A & 4.06 (15.7) & 0.933 (1.32) & 0.295 (0.938) & N/A & N/A & N/A \\
    PC/PS & 0.101 (0.177) & 0.504 (0.413) & N/A & 4.06 (15.9) & 0.830 (1.09) & 0.556 (0.317) & N/A & N/A & N/A \\
    \hline
  \end{tabular}
  \caption{MSD Comparison to True Values, Median (IQR).}
  \label{table:simulation_results_median}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{sim_case5}
  \caption{Simulation Case 5}
  \label{fig:sim_case4}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{adni_lhipp}
  \caption{Left Hippocampus, ADNI Participant}
  \label{fig:adni_result}
\end{figure}
